---
title: "Assignment B-1"
author: "Alex Wang"
output: github_document
---

```{r}
library(tidyverse)
library(testthat)
library(palmerpenguins)
library(datateachr)
library(roxygen2)
```

## Make and Document a Function

```{r}
#' @title 
#' Grouped Count and Proportion
#' 
#' @description
#' 'group_count_prop' groups a dataframe by a variable of your choosing and find the counts and proportions of another variable of your choosing within those groups.
#' 
#' @param df Data frame of interest. Named df since this is short for data frame, which is the class required.
#' @param group_vars Categorical variable present in the data frame. It is used to group the data. Named 'group_vars' since it is the variable by which the data will be grouped.
#' @param var Categorical variable present in the data frame. It is the variable that will be counted and whose proportion will be calculated. Named 'var' since this is the variable that will be counted and whose proportion will be calculated.
#' 
#' @return
#' 'group_count_prop' returns the count and proportion of a categorical variable grouped by another categorical variable in a data frame. The output is a tibble, which has columns corresponding to the count (n) and proportion.

group_count_prop <- function (df, group_vars, var) {
  stopifnot(is.data.frame(df))
  df %>%
    group_by({{group_vars}}, {{var}}) %>%
    summarise(n = n()) %>%
    mutate(proportion = n/sum(n))
}
```

## Examples
### Example using building_permits data set
Using the group_count_prop function, I can group by year and find the counts and proportions by type_of_work in the building_permits data frame. The output tibble contains the year, and type_of_work variables with their corresponding counts (n) and proportions.
```{r}
permits_count_prop <- group_count_prop(building_permits, year, type_of_work)
print(permits_count_prop)
```

### Example using penguins data set 
On a different data frame, called penguins, I can group by species and find the counts and proportions of sex in the data frame. The output tibble contains the species, and sex variables with their corresponding counts (n) and proportions. Since some entries are missing sex, it also displays the count and proportion of missing entries where the sex observation is not given (NA).
```{r}
penguins_count_prop <- group_count_prop(penguins, species, sex)
print(penguins_count_prop)
```

## Test
```{r}
test_that("group_count_prop only accepts data frame", {
  expect_error(group_count_prop(test_vector, year, species)) #when the argument that is not a data frame is used in the function, an error is expected from the function
  expect_error(group_count_prop("building_permits", year, type_of_work)) #when the argument is coerced into a character class, an error is expected from the function
})

test_that("group_count_prop returns the expected number of columns with the expected names", {
  expect_length(group_count_prop(building_permits, year, type_of_work), 4) #when the function is used on a data frame, the expected number of variables in the output tibble is 4 (the variable the data is grouped by, the variable that is counted, the count, and the proportion)
  expect_named(group_count_prop(building_permits, year, type_of_work), c("year", "type_of_work", "n", "proportion")) #the expected names of the columns in the output tibble are 'year', 'type_of_work', 'n', and 'proportion'
})

test_that("group_count_prop proportions equal 1", {
  expect_length(group_count_prop(penguins, species, sex) %>% summarise_at(vars(proportion), 
               list(species = sum)), 1) #the proportions that the function returns should add up to 1 within each group
})

```


