---
title: "Assignment B-1"
output: github_document
---

# Load Relevant Packages
```{r}
library(tidyverse)
library(testthat)
library(palmerpenguins)
library(datateachr)
library(roxygen2)
```

## Make and Document a Function
I want to make a function that bundles a 'group_by() %>% summarise()' workflow. Specifically, my function, named group_count_prop, will take a data frame and group by a categorical variable of interest. Then, it will calculate the number of counts and proportion in each category of another categorical variable. 
```{r}
#' @title 
#' Grouped Count and Proportion
#' 
#' @description
#' 'group_count_prop' groups a data frame by a categorical variable of your choosing and will find the counts and proportions of another categorical variable of your choosing across those groups.
#' 
#' @param df Data frame of interest. Named df since this is short for data frame, which is the class required as an argument for the function.
#' @param group_vars Categorical variable present in the data frame. It is used to group the data. Named 'group_vars' since it is the variable by which the data will be grouped.
#' @param var Categorical variable present in the data frame (different from the group_vars variable). It is the variable that will be counted and whose proportion will be calculated. Named 'var' since this is the variable that will be counted and whose proportion will be calculated.
#' 
#' @return
#' 'group_count_prop' returns the count and proportion of a categorical variable grouped by another categorical variable in a data frame. The output is a tibble, which has columns corresponding to the count (n) and proportion.

group_count_prop <- function (df, group_vars, var) {
  stopifnot(is.data.frame(df)) #stops the function from running if the df argument is not a data frame
  df %>%
    group_by({{group_vars}}, {{var}}) %>% #groups by group_vars variable
    summarise(n = n()) %>% #calculate counts of var variable
    mutate(proportion = n/sum(n)) #calculate proportions
}
```

## Examples
### Example using building_permits data set
Using the group_count_prop function, I can group by year and find the counts and proportions of all different type_of_work categories in the building_permits data frame. The output tibble contains the year, and type_of_work variables with their corresponding counts (n) and proportions.
```{r}
permits_count_prop <- group_count_prop(building_permits, year, type_of_work)
print(permits_count_prop)
```

### Example using penguins data set 
On a different data frame, called penguins, I can group by species and find the counts and proportions of different sex categories in the data frame. The output tibble contains the species, and sex variables with their corresponding counts (n) and proportions. Since some entries are missing sex, it also displays the count and proportion of missing entries where the sex observation is not given (NA).
```{r}
penguins_count_prop <- group_count_prop(penguins, species, sex)
print(penguins_count_prop)
```

## Test the function
Here, I will write formal tests for my function using expect_() functions from the testthat package. I will test to ensure that my function only accepts data frames, that my function returns the expected and numbers of columns in the output tibble, and that my function computes proportions that total to 1.
```{r}
test_that("group_count_prop only accepts data frame", {
  expect_error(group_count_prop(test_vector, year, species)) #when the argument entered in the function is not a data frame, an error is expected from the function
  expect_error(group_count_prop("building_permits", year, type_of_work)) #when the argument is coerced into a character class, an error is expected from the function
})

test_that("group_count_prop returns the expected number of columns with the expected names", {
  expect_length(group_count_prop(building_permits, year, type_of_work), 4) #when the function is used, the expected number of variables in the output tibble is 4 (the variable the data is grouped by, the variable that is counted, the count, and the proportion)
  expect_named(group_count_prop(building_permits, year, type_of_work), c("year", "type_of_work", "n", "proportion")) #the expected names of the columns in the output tibble are 'year', 'type_of_work', 'n', and 'proportion'
})

test_that("group_count_prop proportions equal 1", {
  expect_length(group_count_prop(penguins, species, sex) %>% summarise_at(vars(proportion), 
               list(species = sum)), 1) #the proportions that the function returns should add up to 1 within each group
})

```


